---
title: "draft-functionality-check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{draft-functionality-check}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", cache = TRUE
)
```

```{r setup,warning=TRUE,message=FALSE}
library(pcgen2)
library(ggplot2)
```

# pcgen dependencies:

```{r dependencies}
skipvis <- T
if (!skipvis) {
  # The vanilla dependency graph
  mvbutils::foodweb()

  # The interactive version:
  require(DependenciesGraphs)

  # hacking the envirDependencies() function to get an idea of dependencies:
  name.functions <- as.vector(ls.str( mode = "function"))
  print(name.functions)
  toutfonc <- linksForAll(".GlobalEnv")
  visdata <- prepareToVis(unique(toutfonc), name.functions)
  class(visdata) <- "dependenciesGraphs"
  plot(visdata)

  library(DependenciesGraphs)
  depe <- envirDependencies("package:pcgen2")
  plot(depe, block = T)
  library(pcgen2)
  dep <- funDependencies("package:pcgen2","gencovTest")
  plot(dep)
}

```


## Using Willem's data: ------------------------------------------------------------

This is a dataset from Willem, composed of:
- `d`: a dataframe of 500 genotypes and 2 traits: `d$G`, `d$Y1` and `d$Y2` with 2 replicates per genotype
- `dm`: a dataframe of genotypic marker means and trait means (so it has size: `500*100` )


```{r}
load(file = '~/github_repos/pcgen2/data/simdata1.RData')
str(d)
dm[1:6,1:6]
M <- as.matrix(dm[,-(1:3)] / sqrt(ncol(dm)))
K <- M %*% t(M)

# This evaluates the code 2 times. Increase for a realistic benchmark
mbm <- microbenchmark::microbenchmark(
  "Old w replicates" = {out.old.way.1 <- pcgen::pcgen(d, verbose = T)},
  "New w replicates" = {out.old.way.2 <- pcgen::pcgen(d, verbose = T)},
  "New w/ genric K" = {out.new.way <- pcgen2::pcgen(dm[,1:3], K = K, verbose =T)},
  times = 1)

mbm # Using K is obviously much slower than the other approaches

#autoplot(mbm) +  theme_linedraw() # make times > 2 to see a plot

par(mfrow=c(1,3))
Rgraphviz::plot(out.old.way.1, main = "package output w/ replicates")
Rgraphviz::plot(out.old.way.2, main = "modified code w/replicates")
Rgraphviz::plot(out.new.way, main = "modified code w/ generic K")

all.equal( out.old.way.1, out.old.way.2) # so the new tests didn't break the old code -
                                         # they differ in the call here, which is expected.
all.equal( out.old.way.2, out.new.way)   # so, we learn the same network with and without
                                         # using the kinship matrix. The call and pMax differ
                                         # but this is not surprising


```


# Inner package working

```{r }

y1 <- dm$Y1; y2  <- dm$Y2; G <- dm$G;
## Type A CI tests, verifying the gen.var.test works: ------
# Y1 \perp G (S is empty)
pcgen2:::gen.var.test(y1, K = K) # (No, low p)
# Y1 ~ (1|G) + Y2
pcgen2:::gen.var.test(y1, y2, K = K) # (No, low p)

# Y2 \perp G (S is empty)
pcgen2:::gen.var.test(y2, K=K) # (No, low p)
# Y2 \perp G | Y1
pcgen2:::gen.var.test(y2, y1, K = K) # (Yes, high p)

## Type B tests, verifying res.covar.test works: ---------

Z.t <- pcgen2:::make.Z.matrix(G)
X.t <- Matrix::Matrix(rep(1, length(G)))

em.vec <- c(y1, y2)
names(em.vec) <- rep(as.character(G), 2)

fit.reduced.K <- pcgen2:::fitEM(em.vec, X.t, Z.t, cov.error = TRUE,
                                cov.gen = FALSE, max.iter = 20,
                                K = K)

# But, this seems equivalent to changing Z defition (which is faster)
# in terms of fixed coefficents and variances. Random effects differ.
# The question IS: are they equivalent in terms of likelihood or not?!
# In this case, both models have p-value of 1, so not conclusive
fit.reduced.ZM <- pcgen2:::fitEM(em.vec, X.t, Z.t = Matrix::Matrix(Z.t %*% M),
                                 cov.error = TRUE,
                                 cov.gen = FALSE, max.iter = 20)

all.equal(fit.reduced.K,fit.reduced.ZM) # Difference in random components only.
# It makes sense to investigate this with larger
# number of traits. See section on (Other scenarios)

str(fit.reduced.K)
str(fit.reduced.ZM)

pcgen2:::res.covar.test(y1, y2, G, K = K)  # Y1 \not \perp Y2; so low p expected
# The default was use.manova=T; but that produces an error in this set up (no
# replicates so manova fails). I'm hard coding use.manova=F if K is present.

# It is not sensible that this function returns c(pvalue, TRUE) together. Checking its
# dependencies, TRUE is not used by any other function and is always ignored. Maybe just
# remove it here?

# Also, when calling fitEM function for the reduced model, I'm setting cov.error=T;
# cov.gen=F as per the examples above; but I don't really understand these options

## Calling pcgenTest -------------------------------------------------------------------

suffStat <- dm[,1:3]; K <- K
## case 1; Trait \perp Trait | {Traits, G, Q}
x <- 2; y <- 3; S <- 1;
pcgen2:::pcgenTest(x, y, S, suffStat, K = K) # Y1 \perp Y2 | G (No, low p)

## case 2; Trait \perp G | Traits
x <- 2; y <- 1; S <- 3;
pcgen2:::pcgenTest(x, y, S, suffStat, K = K) # Y1 \pepr G | Y2 (No, low p)

x <- 3; y <- 1; S <- 2;
pcgen2:::pcgenTest(x, y, S, suffStat, K = K) # Y2 \pepr G | Y1 (Yes, high p)

## case 3: Trait \perp QTL | {Trait, QTL, G}
# Needs further investigation


## Calling skeleton2 -------------------------------------------------------------------
skel <- pcgen2:::skeleton2(suffStat, alpha = .01, K = K, labels = colnames(suffStat), verbose = T)
Rgraphviz::plot(skel)


out.new.way <- pcgen2:::pcgen(suffStat, K = K, verbose = T); Rgraphviz::plot(out.new.way)

```

## Other scenarios: --------------------------------------------------------------
## 1. The effect of K:
# giving pcgen an uninformative K matrix; or one that is calculated differently
## 2. The presence of more traits

```{r}
# Uninformative Kinship matrix
out.diag.K <- pcgen2::pcgen(dm[,1:3], K = diag(1, 500, 500), verbose = T, NAdelete = F)
Rgraphviz::plot(out.diag.K) # all edges lost

gy <- pcgen2::pcRes(dm[,1:3], K = diag(1, 500, 500))
Rgraphviz::plot(gy) # However, pcRes is able to predict the Y1 --> Y2 edge

# also, pcgen if given residuals:
cor <- cor(pcgen2::getResiduals(suffStat = dm[,1:3], K = diag(1, 500, 500)))
gy2 <- pcgen2::pcgen(dm[,1:3], use.res = T, K = diag(1, 500, 500), res.cor = cor, verbose = T,
             NAdelete = F, return.pvalues = T)
gy2$pMax
Rgraphviz::plot(gy2$gr) # Giving the correct K (to pcgen adds the G-->Y1 edge and orients the Y1-->Y2
          # edge. An uninformative K loses G--> Y1 edge and the direction on Y1 --> Y2
          # This is somewhat not surprising

## How about other definitions of K?
# Generate the raw  0,2 dosage values (for Gaston)
genos <- as.matrix(purrr::map_df(purrr::map_df(dm[, -c(1:3)], factor), as.integer))
genos[genos == 1] <- 0
rownames(genos) <- rownames(dm)
genos.gaston <- as(genos,"bed.matrix")
K2 <- gaston::GRM(genos.gaston, autosome.only = F) # calculated as: XX'/q, with X the standardized
                                           # genotype matrix and q the number of SNPs (ncol(x))

all.equal(K, K2)  # mean difference of ~1

out.new.K <- pcgen2::pcgen(dm[,1:3], K = K2, verbose = T, NAdelete = T)
Rgraphviz::plot(out.new.K)  # Only the Y1 -- Y2 edge remains but with no orientation

cor <- cor(pcgen2::getResiduals(suffStat = dm[,1:3], K = K2))
gy3 <- pcgen2::pcgen(dm[,1:3], use.res = T, K = K2, res.cor = cor)
Rgraphviz::plot(gy3)        # We still lose the G-->Y1 edge

# How about selecting the very commong snps for defining K?
genos.trimmed <- gaston::select.snps(genos.gaston, maf > .2) # 70 snps
K3 <- gaston::GRM(genos.trimmed, autosome.only = F)
out.new.trimmed <- pcgen2::pcgen(dm[,1:3], K = K3, verbose = T, NAdelete = T)
Rgraphviz::plot(out.new.trimmed)  # This gives the correct network

# and if selecting the rare snps:
genos.trimmed2 <- gaston::select.snps(genos.gaston, maf < .2) # 30 snps
K4 <- gaston::GRM(genos.trimmed2, autosome.only = F)
out.new.trimmed2 <- pcgen2::pcgen(dm[,1:3], K = K4, verbose = T)
Rgraphviz::plot(out.new.trimmed2)  # We lose the G-->Y1 edge

out.new.trimmed2 <- pcgen2::pcgen(dm[,1:3], K = K4, verbose = T, NAdelete = T)
Rgraphviz::plot(out.new.trimmed2)  # We lose the G-->Y1 edge

## How about having more traits?
suffStat <- cbind(dm[,1:3], Y3 = rnorm(n = nrow(dm)),
                  Y4 = dm$Y1 + runif(n = nrow(dm)))
out.more.traits <- pcgen2::pcgen(suffStat, K = K, verbose = T)
Rgraphviz::plot(out.more.traits)

cor2 <- cor(pcgen2::getResiduals(suffStat = suffStat, K = K))
gc2 <- pcgen(suffStat, use.res = T, K = K, res.cor = cor2)
Rgraphviz::plot(gc2) # The G --> Y1 edge is lost; and logically non directed edges otherwise

out.pc.res <- pcgen2::pcRes(suffStat, K = diag(1, 500, 500))
Rgraphviz::plot(out.pc.res)
```

# Futuristic routes: -----------------------------------------------------

1. Borrowing some functionality from GCTA.
    - for calculating the GRM for example, GCTA has fast way of doing
    - for finding the variance components, GCTA also has nice features
    - The recent fastGWA way also looks cool --> see how a function like: 
  `OmicKriging::computeGRM(bedfullheader, grmfullheader = "genomic", gctaname = "gcta", idfile = NULL, snpfile = NULL)`

2. Consider adopting functionality and methods from `gwasTools`.
    - Has efficient means of storing genomic data in general
    - Also, methods for QC and relevant plots .. etc are readily available
