---
title: "Complete pcRes pcgen pipeline"
author: "Azza Ahmed"
date: "7/2/2021"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
pacman::p_load(tidyverse, cowplot)
devtools::load_all(path = "/home/azza/github_repos/pcgen2/")
```

## Simulate data

We are simulating data with traits models tha tfollows:

$Y = XB +  \lambda Y + G + E$, where 
   $G \sim MVN (0, K, \Sigma_G): vec(G) \sim N (0,  \Sigma_G \bigotimes K)$
   $E \sim MVN (0, I_n, \Sigma_E): vec(E) \sim N(0,  \Sigma_E\bigotimes I_n )$

Or, by rearranging:
$Y = XB\Gamma +  G\Gamma +  E\Gamma$
$vec(Y) \sim N(vec(XB), \Gamma^t\Sigma_G\Gamma \bigotimes K + \Gamma^t\Sigma_E\Gamma \bigotimes I_n) \sim N(vec(XB), V_G \bigotimes K + V_E \bigotimes I_n)$


```{r}
library(GraphPhenoSimulator)
load("~/github_repos/GraphPhenoSimulator/data/simdata1.RData")
X <- as.matrix(dm[, -c(1:3)])
K <- X %*% t(X) / sqrt(ncol(X))
rm(list = c("d", "dm", "X"))
K <- K + 0.005 *diag(nrow(K))
ng <- nrow(K)    # number of genotypes
reps <- 1                # number of (genetically identical) replicates of each genotype
randomizeSamples <- F # reorder the kinship matrix rows and columns?
negative <- FALSE     # Create a DAG with negative edge weights?
topsort <- T          # If TRUE, non-genetic traits are chosen from nodes with highest topological order
fixngen <- TRUE       # If TRUE, Number of traits with 0 genetic effects = ((1-pgen) * ntraits).
# If FALSE, the number is random with probability = 1 - pgen
useK <- TRUE          # use a marker-based kinship matrix
simulateQTLs <- T     # simulate QTLs, explaining part of the genetic variance ?
alphaQTL <- 0         # The proportion of genetic variance explained by the QTLs
fixnumber <- T        # If TRUE, Number of QTLs = lambdaQTL
# If FALSE, number is drawn from Poisson with lambdaQTL
lambdaQTL <- 2        # Number of QTLs or the Poisson parameter for their number
pQTL <- 0.5           # The probability a QTL has an effect on a trait
noise <- F       # Add noise to the simulated graph data?
extranoise <- 0.2 # Additional noise amount




## Are these values standard? Or are they arbitrary..
ntraits <- 3         # number of traits
dagcoeff <- c(1, 1) # Lower and upper limits of DAG edge weights -------------> try with 1 edge, change its weight
######################################### Is it the strength of correlation? when it is large/small
######################################### Have an ideal data with strong relations between the data
rangeGenVar <- c(.1,.2) # Range of variances of direct genetic effects  ---------------> make it low
pgen <- 10           # Probability a trait has non-zero genetic variance.
rangeEnvVar <- c(.1,.2) # Range of error variances (environment effects) ------------------> make it low


if (randomizeSamples) {
  genotype.sample <-  sample(1:ng, size = ng) # if looking for randomness!
  K <- K[genotype.sample, genotype.sample]
}
p.edge <- 1 * min(1, 1/(ntraits-1)) # probability of an edge, in the randomDAG
```

## Learn the effects of changing parameters

1. how are the learnt networks different?

```{r}
runs <- 8
trait.data <- list()
bench.dag <- tibble::tibble(tpr = numeric(runs), fpr = numeric(runs),
                            tdr = numeric(runs), shd = numeric(runs))

bench.dag <- cbind(matrix(0, ncol = 3, nrow = runs), bench.dag)
names(bench.dag)[1:3] <- c("dagcoeff", "rangeGenVar", "rangeEnvVar")

bench.gdag <- tibble::tibble(tpr = numeric(runs), fpr = numeric(runs),
                             tdr = numeric(runs))

counter <- 1
for (dagcoeff in c(1, 10)){
  for (rangeGenVar in c(.1, 1))
    for (rangeEnvVar in c(.1, 1)){
      # generate data:
      print(counter)
      dag <- matrix(0, ntraits, ntraits)
      
      dag[1, 2] <- dagcoeff
      rownames(dag) <- colnames(dag) <- paste("Trait", 1: ntraits)
    
      
      geneticCovMatrix <- clusterGeneration::genPositiveDefMat(covMethod = "onion",
                                                               dim=ntraits, alphad=1,
                                                               rangeVar=c(rangeGenVar, rangeGenVar))$Sigma
      rownames(geneticCovMatrix) <- colnames(geneticCovMatrix) <- colnames(dag)
      
      traits.without.direct.gen.effect <- rownames(dag)[3]
      
      geneticCovMatrix[traits.without.direct.gen.effect, ] <- 0
      geneticCovMatrix[ ,traits.without.direct.gen.effect] <- 0
      
      errorVars <- runif(min = rangeEnvVar[1], max = rangeEnvVar[1], n = ntraits)
      names(errorVars) <- names(dag)
      
      gsem <- create.gsem(dag.adj = dag, genCovMatrix = geneticCovMatrix,
                          errorVars = errorVars)
      Rgraphviz::plot(gsem[['gdag']], main = "Simulated  genetic network")
      
      test <- rmvDAG_H2(ng = ng, r = r, gsem = gsem, simulate.QTLs = simulateQTLs,
                        lambda.QTL = lambdaQTL, fix.number = fixnumber,
                        alpha.QTL = alphaQTL, p.QTL = pQTL,
                        K = K, use.K.names = TRUE, use.marker.matrix = T, M = M)
      
      nw.data <- test$fulldata[, 1:(ntraits+1)]
      
      trait.data[[counter]] <- cor(nw.data[, -1])["Trait.1", "Trait.2"]
      
      learnt.dag <- pcRes(nw.data, K = K,
                          cov.method = "uni", m.max = Inf, verbose = F, alpha = 1e-2)
      plot.nw(learnt.dag, "pcRes learnt network", fontsize = 20)
      
      cor <- cor(getResiduals(nw.data, cov.method = "uni", K = K))
      learnt.gdag <- pcgen(nw.data, K = K, res.cor = cor, use.res = T, verbose = F, alpha = 1e-2)
      plot.nw(learnt.gdag, "pcgen learnt network", fontsize = 20)
      
      bench.dag[counter, c("tpr", "fpr", "tdr")] <- t(pcalg::compareGraphs(learnt.dag@graph, gsem$dag))
      bench.dag[counter, "shd"] <- pcalg::shd(learnt.dag@graph, gsem$dag)
      bench.dag[counter, 1:3] <- c(dagcoeff, rangeGenVar, rangeEnvVar)
      
      plot.nw(learnt.dag, "pcRes learnt network", fontsize = 25)
      
      bench.gdag[counter, c("tpr", "fpr", "tdr")] <-
        t(GraphPhenoSimulator::compareGeneticEffects(learnt.gdag@graph, gsem$gdag))
      counter <- counter + 1
    }
}

```


```{r}
bench.dag <- cbind(matrix(0, ncol = 3, nrow = 6), bench.dag)
names(bench.dag)[1:3] <- c("dagcoeff", "rangeGenVar", "rangeEnvVar")
counter <- 1

for (dagcoeff in c(.1, .5, 1, 5, 10))
  for (rangeGenVar in c(.01, .1, 1, 10))
    for (rangeEnvVar in c(.01, .1, 1, 10)){
      bench.dag[counter, 1:3] <- c(dagcoeff, rangeGenVar, rangeEnvVar)
      counter <- counter + 1
    }

```


```{r plotParameterEffects}
g1 <- bench.dag %>% pivot_longer(-shd, names_to = "metric", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = metric, y = value)) +
  geom_boxplot(aes(x = metric, y = value)) + theme_bw() +
  labs(title = "Comparing traints-only network")

g2 <- bench.dag %>% pivot_longer(shd, names_to = "metric", values_to = "value") %>%
  ggplot(aes(x = metric, y = value)) +
  geom_point() + geom_boxplot() + theme_bw() + theme(axis.title.y = element_blank())

g3 <- bench.gdag %>% pivot_longer(everything(), names_to = "metric", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = metric, y = value)) +
  geom_boxplot(aes(x = metric, y = value)) + theme_bw() +
  labs(title = "Comparing genetic effects only")



plot_grid( g1, g2 , g3, rel_widths = c(2, 1, 2))
```



## Now, test with strong network

```{r tstParameterMeaning}
runs <- 10

bench.dag <- tibble::tibble(tpr = numeric(runs), fpr = numeric(runs),
                            tdr = numeric(runs), shd = numeric(runs))

bench.gdag <- tibble::tibble(tpr = numeric(runs), fpr = numeric(runs),
                             tdr = numeric(runs))
for (i in 1:runs){
  print("---------------------------------------------------------------------")
  print(i)
  print("---------------------------------------------------------------------")
  
  ## Generate the data
  { dag <- randomDAG2(n = ntraits, prob = p.edge, lB = dagcoeff[1], uB = dagcoeff[2],
                      V = paste0('Trait', 1:ntraits), negative = negative)
    Rgraphviz::plot(dag, main =
                      paste("Simulated Traits only network- Erdoes-Renyi: p.edge =",
                            p.edge, "weights range = ", dagcoeff[1], dagcoeff[2]))
    
    geneticCovMatrix <- clusterGeneration::genPositiveDefMat(covMethod = "onion",
                                                             dim=ntraits, alphad=1,
                                                             rangeVar=rangeGenVar)$Sigma
    rownames(geneticCovMatrix) <- colnames(geneticCovMatrix) <- graph::nodes(dag)
    
    if (pgen < 1) {
      if (fixngen == TRUE) {
        if (topsort == TRUE) {
          traits.without.direct.gen.effect <-
            rev(match(names(igraph::topological.sort(igraph::graph_from_graphnel (dag))),
                      names(graph::edgeL(dag)))) [1:(round((1 - pgen) * ntraits))]
        } else  # topsort == FALSE
          traits.without.direct.gen.effect <-
            sample(1:ntraits, size = round((1 - pgen)*ntraits))
      } else  # fixngen == F
        traits.without.direct.gen.effect <-
          which(rbinom(p = 1 - pgen, size=1, n = ntraits) == 1)
    } else  # pgen >=1
      traits.without.direct.gen.effect <- integer()
    
    geneticCovMatrix[traits.without.direct.gen.effect, ] <- 0
    geneticCovMatrix[ ,traits.without.direct.gen.effect] <- 0
    
    errorVars <- runif(min = rangeEnvVar[1], max = rangeEnvVar[2], n = ntraits)
    names(errorVars) <- graph::nodes(dag)
    
    gsem <- create.gsem(dag.adj = dag, genCovMatrix = geneticCovMatrix,
                        errorVars = errorVars)
    Rgraphviz::plot(gsem[['gdag']], main = "Simulated  genetic network")
    
    test <- rmvDAG_H2(ng = ng, r = r, gsem = gsem, simulate.QTLs = simulateQTLs,
                      lambda.QTL = lambdaQTL, fix.number = fixnumber,
                      alpha.QTL = alphaQTL, p.QTL = pQTL,
                      K = K, use.K.names = TRUE, use.marker.matrix = T, M = M)
    
    nw.data <- test$fulldata[, 1:(ntraits+1)]
    if (noise == T) {
      for (j in (2:(ntraits+1)))
        nw.data[,j] <- nw.data[,j] + rnorm(n = nrow(d),
                                           sd = sqrt(extranoise * gsem$Ve[j-1,j-1]))
    }
  }
  
  ## Learn the networks
  learnt.dag <- pcRes(nw.data, K = K,
                      cov.method = "uni", m.max = Inf, verbose = F, alpha = 1e-3)
  
  cor <- cor(getResiduals(nw.data, cov.method = "uni", K = K))
  
  learnt.gdag <- pcgen(nw.data, K = K, res.cor = cor, use.res = T, verbose = F)
  
  ## Evaluate the learning:
  
  bench.dag[i, c("tpr", "fpr", "tdr")] <- t(pcalg::compareGraphs(learnt.dag@graph, gsem$dag))
  bench.dag[i, "shd"] <- pcalg::shd(learnt.dag@graph, gsem$dag)
  
  
  bench.gdag[i, c("tpr", "fpr", "tdr")] <-
    t(GraphPhenoSimulator::compareGeneticEffects(learnt.gdag@graph, gsem$gdag))
}
```







## Learning with pcRes

```{r}

par(mfrow=c(1,3))
plot.nw(gsem$gdag, "True Genetic and traits network", fontsize = 20)
plot.nw(gsem$dag, "True traits-only network", fontsize = 20)
plot.nw(learnt.dag, "pcRes learnt network", fontsize = 25)


# eval.nw(runs = 5, gsem$dag, suffStat = nw.data,
#         K = K, cov.method = "uni", m.max = Inf)

```
## Learning with pcgen

```{r}

par(mfrow=c(2, 2))
plot.nw(gsem$gdag, "True Genetic and traits network", fontsize = 20)
plot.nw(gsem$dag, "True traits-only network", fontsize = 20)

plot.nw(learnt.dag, "pcRes learnt network", fontsize = 25)

```

