---
title: "Tuning simulaiton parameters"
author: "Azza Ahmed"
date: "7/22/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(dscrutils, tidyverse)
```

## Set up

We are trying to see the effect of the simulation parameters in the learnt network, 
i.e, which combination of parameters lead `pcgen` and `pcRes` to infer the correct network
So, we:
1. test with parameter combinations: 
    * `dagcoeff`: 0.01, 0.1, .5, 1, 5
    * `rangeGenVar`: c(.1, .2), c(1,2), c(4,5), c(9,10) # try using the same value lb and up; the tendency should be larger range, results in larger genetic effects.
    * `rangeEnvVar`: c(.1, .2), c(1,2), c(4,5), c(9,10)
2. learn without genotype


## Simulation 1:

We simulate a simple 3 node network using the script: `simulation_tuning.dsc`.
Here, *Trait1* causes *Trait2* (with `dagcoeff`), and both are genetic, while 
*Trait3* is indepenent of all.

The expectation here is that:
1. `pcRes` should learn the skeleton: $Trait1 \longleftrightarrow Trait2$
2. `pcgen` should learn the skeleton: $G \rightarrow Trait1; G \rightarrow Trait2; Trait1 \longleftrightarrow Trait2 $


```{r}

dscout <- dscquery(dsc.outdir= "progress/simulation_tuning/", 
                   targets = c("simulate.dagcoeff", "simulate.rangeGenVar", 
                               "simulate.rangeEnvVar", "simulate.true_dag",
                               "simulate.data", "simulate.cor_traits",
                               "learn", "learn.learnt_dag", "learn.cor_residuals",
                               "learn.dagfit_tpr", "learn.dagfit_fpr", 
                               "learn.dagfit_tdr", "learn.dagfit_shd", 
                               "learn.genfit_tpr", "learn.genfit_fpr", 
                               "learn.genfit_tdr" ))  %>% as_tibble()
```

### Zoom-in to 1 simulation
So, we investigate a bit the networks visually, before looking into summary stats:

```{r}
nw.pc.vanilla <- pcalg::pc(suffStat = list(C = cor(dscout$simulate.data[[1]][, -1]), 
                                           n = nrow(dscout$simulate.data[[1]])),
                       indepTest = pcalg::gaussCItest, alpha = 0.01,
                        labels = names(dscout$simulate.data[[1]][,-1]), m.max = Inf, # for pcRes
                       skel.method = "stable", u2pd = "relaxed", NAdelete = FALSE,
                       conservative = FALSE, maj.rule = TRUE, solve.confl = TRUE)

par(mfrow = c(2, 3))

plot.nw(dscout$simulate.true_dag[[2]]$dag, "Simulated traits-only network", fontsize = 20)
plot.nw(dscout[dscout$learn == "learntraits", ]$learn.learnt_dag[[1]], "pcRes learnt network", fontsize = 20)
plot.nw(nw.pc.vanilla, "pcalg::pc network", fontsize = 8)


plot.nw(dscout$simulate.true_dag[[2]]$gdag, "Complete simulated  Genetic network", fontsize = 20)
plot.nw(dscout[dscout$learn == "learngtraits", ]$learn.learnt_dag[[1]], "pcgen learnt network", fontsize = 20)

```


### Residuals

Q: Would vanilla `pc` act similarly to `pcRes`?
Guess: This depends on the relation between traits vs residulas

```{r}
dscout %>% filter(learn == "learngtraits") %>% 
  select(simulate.cor_traits, DSC, learn.cor_residuals) %>% 
  mutate(cor_residuals = learn.cor_residuals %>% map_dbl(3),
         cor_traits = simulate.cor_traits,
         replicate = factor(DSC)) %>%
  ggplot(aes(x = cor_traits, y = cor_residuals, group = replicate)) + geom_point(aes(, color = replicate)) +
  geom_smooth(
               se = FALSE) +
  theme_classic()
  
```


### Aggregate results

```{r }
dscout <- dscout %>% 
  group_by(simulate.dagcoeff,simulate.rangeGenVar,simulate.rangeEnvVar, learn) %>% 
  summarise(across(learn.dagfit_tpr:learn.genfit_tdr, mean)) 

dscout %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_tpr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_fpr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_tdr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_shd, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% filter(learn == "learngtraits") %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_tpr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% filter(learn == "learngtraits") %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_fpr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()

dscout %>% filter(learn == "learngtraits") %>% ggplot() + 
  geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_tdr, 
                  color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, width = 0.02, alpha = .6)+  theme_classic()


dscout %>% filter(learn == "learngtraits") %>% ggplot() + 
  geom_contour_filled(aes( z = learn.genfit_tdr, 
                  x = as.numeric(as.factor(simulate.rangeGenVar)), y = as.numeric(as.factor(simulate.rangeEnvVar)))
               )+  theme_classic()

```


## Simulation 2:

This is a simple 3 nodes network where *Trait1* causes *Trait2* (with `dagcoeff`),
while only *Trait3* has direct genetic effect. The simulation script used here is:
`simulation_tuning_3genetic.dsc`

The expectation here is that:
1. `pcRes` should learn the skeleton: $Trait1 \longleftrightarrow Trait2$
2. `pcgen` should learn the skeleton: $G \rightarrow Trait3;  Trait1 \longleftrightarrow Trait2 $

```{r}
pacman::p_load(dscrutils, tidyverse)

dscout_3g_simplified <- dscquery(dsc.outdir= "progress/simulation_tuning_3genetic/", 
                   targets = c("simulate.dagcoeff", "simulate.rangeGenVar", 
                               "simulate.rangeEnvVar", "learn",
                               "learn.dagfit_tpr", "learn.dagfit_fpr", 
                               "learn.dagfit_tdr", "learn.dagfit_shd", 
                               "learn.genfit_tpr", "learn.genfit_fpr", 
                               "learn.genfit_tdr", 
                                "simulate.cor_traits", "learn.cor_residuals",
                               "simulate.true_dag", "learn.learnt_dag", "learn.learnt_gdag",
                               "simulate.genvar"))  

dscout_3g_simplified$simulate.dagcoeff[[1]]
plot.nw(dscout_3g_simplified$learn.learnt_dag[[1]], "DAG", fontsize = 10)

dscout_3g_simplified$simulate.dagcoeff[[4]]
dscout_3g_simplified$simulate.genvar[[4]]
plot.nw(dscout_3g_simplified$learn.learnt_gdag[[3]], "DAG", fontsize = 10)


index <- which(dscout_3g$simulate.dagcoeff == .1)
dscout_3g$simulate.cor_traits[index]

dscout_3g$learn.dagfit_tpr[[1]]
C=dscout_3g_simplified$learn.cor_residuals[4]

dscout_3g$simulate.rangeGenVar[index][50]
dscout_3g$simulate.rangeEnvVar[index][50]
Rgraphviz::plot(pcalg::pc(suffStat = list(C = C, n = 500), labels = rownames(C), 
                          indepTest = gaussCItest, alpha = .01))


pdf("jitters_3gen.pdf")
dscout_3g %>% ggplot() + geom_point(aes(x = simulate.dagcoeff, y = learn.dagfit_tpr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar)
                                  )+ theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_fpr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_tdr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.dagfit_shd, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_tpr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_fpr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()

dscout_3g %>% ggplot() + geom_jitter(aes(x = simulate.dagcoeff, y = learn.genfit_tdr, 
                                      color = simulate.rangeGenVar, shape = simulate.rangeEnvVar),
              height = .02, alpha = .6)+  theme_classic()
dev.off()



pdf("lines_3gen.pdf")
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.dagfit_tpr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.dagfit_fpr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.dagfit_tdr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.dagfit_shd, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.genfit_tpr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.genfit_fpr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dscout_3g %>% ggplot() + geom_line(aes(x = simulate.dagcoeff, y = learn.genfit_tdr, 
                                    color = simulate.rangeGenVar, linetype = simulate.rangeEnvVar)) + theme_classic()
dev.off()


```


